---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The zip file "activity.zip" should be in the working directory.

Unzip the data file and read "activity.csv" and convert interval column into a factor

Examine the first few rows of data

```{r}
unzip("activity.zip")
data <- read.csv(file="activity.csv", colClasses=c("numeric", "Date", "factor"))
head(data)
```


## What is mean total number of steps taken per day?
Calculate aggregate of total steps per day and plot the results into a histogram

```{r}
dataDailySteps <- aggregate(data$steps, list(data$date), sum)
hist(dataDailySteps$x)
```

Calculate the mean of number of daily steps
```{r}
mean(dataDailySteps$x, na.rm = TRUE)
```

Calculate the median of the number of daily steps
```{r}
median(dataDailySteps$x, na.rm = TRUE)
```


## What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days
```{r}
dataMeanDailySteps <- aggregate(data$steps, list(data$interval), mean, na.rm=TRUE)
with(dataMeanDailySteps, plot(Group.1, x, type='l', xlab='Interval', ylab='Steps'))
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
dataMeanDailySteps[which.max(dataMeanDailySteps$x),]$Group.1
```

## Inputing missing values



## Are there differences in activity patterns between weekdays and weekends?
